---
title: "Budweiser Data Analysis"
author: "Laura Ahumada"
date: "10/5/2021"
output: html_document
---
  



```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r include=FALSE}
#Load Libraries
library(gridExtra)
library(dplyr)
library(tidyr)
library(stringi)
library(rvest) #html_table, html_node
library(ggplot2)
library(RCurl) #getURL
library(naniar)
library(e1071)
library(tm) 
library(stringr)
library("kableExtra")
library(ggthemes)
library("ggpubr")
library(class)
library(caret)
library(usmap) #import the package
library(ggplot2) #use ggplot2 to add layer for visualization

```


- Here are the answer to your questions and some extra analysis that might be useful

```{r include=TRUE}
#Reading Files
#Open breweries file
breweries=read.csv("/Users/lauraahumada/Documents/SMU_main/DoingDataScience/MSDS_6306_Doing-Data-Science/Unit 8/Breweries.csv")
#Open Beers file
beers=read.csv("/Users/lauraahumada/Documents/SMU_main/DoingDataScience/MSDS_6306_Doing-Data-Science/Unit 8/Beers.csv")
```

<!-- Obtaining the amount of breweries per State in descending order. -->
```{r}
all=breweries %>% group_by(State) %>% summarize(Count=n()) #Breweries per State
all$State= factor(all$State, level =all$State[order(all$Count)]) #Organize by State
```

```{r out.width="100%", out.height="150%"}
#Graph
ggplot(all, aes(x = as.factor(State), y = Count, color=State)) +
  geom_bar(stat = "identity", position = "dodge", width=.5) +
  geom_text(size=2.5,position = position_dodge(width= .5), aes(label=Count),vjust= .5, hjust = -0.5)+
  ggtitle("Breweries per State") + 
  ylab("Number of Breweries") +
  xlab("State") +
  coord_flip() +
  theme(plot.title = element_text(hjust = 0.5),legend.position = "none") +
  theme(axis.text = element_text(size = 6)) # changes axis labels
```

- There is a range of 1-47 Breweries per State\
- The state with most Breweries is Colorado\
- The state with the least Breweries is DC\
- Odd to see Texas, the second largest state in USA, as the 5th State with most Breweries. One would imagine they would be number 1 or 2nd with most breweries.\


# Calculating the missing values
```{r}

#Total missing values from both data frames
all_missing=(data_frame(Count=(c(sapply(beers, function(x) all_miss=sum(is.na(x))),sapply(breweries, function(x) sum(is.na(x))))), columns=c((names(beers)),(names(breweries)))))

#Verify if there are no missing char values 
#Find Empty string in the char columns
#summary(breweries) #Name, city and State are characters columns except the ID 
name=breweries %>% filter((Name=="")|(Name=="NA")|(Name==" ")) #None
city=breweries %>% filter((City=="")|(City=="NA")|(City==" ")) #None
stt=breweries %>% filter((State=="")|(State=="NA")|(State==" ")) #None

#summary(beers) #Style and Name columns are characters rest are numbers
#Find Empty string in the char columns
styl=beers %>% filter((Name=="")|(Name=="NA")|(Name==" ")) #None
missing_style=beers %>% filter((Style=="")|(Style=="NA")|(Style==" ")) #5 NA
all_missing[6,"Count"]=5 #Update missing value for Style in Data frame 

```

```{r, fig.width=11}
#Visualizing missing values
all_missing %>% ggplot(aes(x=Count, y=columns, color=columns)) + geom_col(position = "dodge") +
  xlab("Total amount of missing values") + ggtitle("Missing values per variable") +
  geom_text(size=3,position = position_dodge(width= .5), aes(label=Count),vjust= .5, hjust = -0.5) +
  theme(axis.text = element_text(size = 10)) + ylab("Variable")

```

### The missing values: \
- International Bitterness Units of the beer has 1005 missing values \
  - That is half the data. If allowed based on further analysis on data, values should be imputed \
- Alcohol by volume of the beer has 62 missing values \
- Bitternes unit has very few data missing, it can be ommited.
- Style has 5 missing values \
- Name of beer, Beer Id, Brewery ID, Ounces, Brew_ID, Name of Brewery and City have no values missing \



```{r include=FALSE}
#Looking at the distributions and correlations
library(GGally)
test2=(beers[,c(2,3,4,5,7)])
#All variable with their distributions and correlation
#Find attribute to fill missing values 
ggpairs(test2, aes(alpha = 0.4)) #IBU and ABV suggested a pattern
#The correlation between IBU and ABV is .67 per Pairplot
```


- Checking correlation to see if IBU NA data can be imputed

```{r message=FALSE}
#ABV can help fill 943 IBU missing values since pairplot shows that there is a 67% correlation

#Visualizing IBU and ABV pattern to confirm
#ABV vs IBU in segments in box plot to see distribution
beers %>% 
ggplot(aes(x= ABV, y=IBU, color=as.factor(cut_width(beers$ABV,.009)))) + 
geom_boxplot(mapping = aes(group = cut_width(ABV,.009))) +
  ggtitle("Alcohol volumne vs Bitterness unit") + xlab("Alcohol by volume") +
  ylab("Bitterness unit") +
  scale_colour_discrete(name="Alcohol Segment") 
```

- The mean keeps increasing, suggesting a pattern

```{r include=FALSE}
# Visualizing it

#Plots used to for getting see distriution of each of the Segmentation of the ABV and IBV for missing values
c=ggplot(data = beers, mapping = aes(y = ABV)) +
  geom_boxplot(color = "#FF33CC") +
  ggtitle("Alcohol by volume") + ylab("Alcohol by volume") 
c
```
```{r include=FALSE}
#ggplotly(c)
c=ggplot(data = beers, mapping = aes(y = IBU)) +
  geom_boxplot(color = "#00AFBB") +
  ggtitle("International bitterness unit") + ylab("Bitterness") 
c
#ggplotly(c)

```

- IBU mean per segment using ABU was used to fill NA's and dropped the ABU missing values

```{r include=FALSE}
# Treating missing values
new=beers #New Data frame to manipulate

#Function to split Quartile
#These will be the break limits
quartile_division<- function(qt) {
  n=length(qt)
  mid_n=c()
  for (i in c(1:n)){
    mid_n=c(mid_n,qt[i])
    if (i!=n) {
      mid=(((qt[i+1]-qt[i])/2)+qt[i])
      mid_n=c(mid_n, mid)
      #mid_n=c(mid_n, median(qt[i:i+1]))
    }
  }
  return(mid_n)
}

#Function to get mean per section
quartile_mean<- function(qt) {
  n=length(qt)
  mean_n=c()
  for (i in c(1:n)){
    if (i!=n) {
      mid=(((qt[i+1]-qt[i])/2)+qt[i])
      mean_n=c(mean_n,mid)
      #mid_n=c(mid_n, median(qt[i:i+1]))
    }
  }
  return(mean_n)
}



#Get Segments
qt1=summary(beers$IBU) #Getting the Quartiles
qt=c(qt1[1:3],qt1[5],127) #Removing the Mean and adding upper fence
IBU_s=c(quartile_division(qt),qt1[6]) #IBU distribution in 8 segment 
av_ibu=c(quartile_mean(IBU_s)) #IBU distribution mean per segment


#New IBU mean column according to the IBU value
new$IBU_mean=cut(new$IBU, breaks=IBU_s,labels = av_ibu) 
#Categories from 1-9
#New IBU label column according to IBU value
new$IBU_ind=cut(new$IBU, breaks=IBU_s,labels = seq(1, 9,1)) 

qt=summary(new$ABV) #Getting the Quartiles

qt=c(qt[1:3],qt[5],0.091) #Removing the Mean and adding the upper fence
ABV_se=c(quartile_division(qt),0.14) #ABV distribution separated in 8 segments
av_abv=c(quartile_mean(ABV_se)) #ABV distribution mean per segment


#New ABV mean column based on ABV value
new$ABV_mean=cut(new$ABV, breaks=ABV_se,labels = av_abv) 
#Categoies from 1-9
#New ABV label column based on ABV value
new$ABV_ind=cut(new$ABV, breaks=ABV_se,labels = seq(1,9,1)) 


#categories for both 9, 9 levels!!
#these categories also correspond with the index to can point to the mean
length(av_ibu)
length(av_abv)


#if IBU is NA,then use ABV_num to replace with mean of IBU
new$IBU=ifelse ((is.na(new$IBU) & !is.na(new$ABV)),av_ibu[new$ABV_ind],new$IBU)
#Remove columns created to fill missing values
new=new %>% filter(!is.na(IBU)) %>% select(-IBU_mean,-IBU_ind,-ABV_mean) 
#Only 63 values should be dropped (the reows with ABV values missing)
sapply(new, function(x) sum(is.na(x))) #Verify no missing values 
```


# Merging brewery and beer data
```{r}
all_data=merge(breweries, new, by.x = "Brew_ID", by.y = "Brewery_id")
#Verifying that there are no NA's
sapply(all_data, function(x) sum(is.na(x))) #No NA's
```
- First 6 observations of all the data
```{r kable, results = 'asis' }
kable(head(all_data,6),"simple")

```

- Last 6 observations of all the data
```{r kable2, results = 'asis' }
kable(tail(all_data,6),"simple")
```


# Calculating bitterness and alcohol mean per State

```{r}
#IBU mean per State
ibu_state=all_data %>% group_by(State) %>% summarize(IBU_mean=mean(IBU))
abv_state=all_data %>% group_by(State) %>% summarize(ABV_mean=mean(ABV)) 
#Combine both ABV and IBU count per state
IBU_ABV=merge(abv_state,ibu_state, by="State", all.y=TRUE) 
#intersect((head((IBU_ABV %>% arrange(desc(ABV_mean))  %>% select(State)),10)),(head((IBU_ABV %>% arrange(desc(IBU_mean)) %>% select(State)),10))) #Common states in the top 10 for both IBU and ABV

```
```{r, fig.width=8,fig.height=7}
#Bitterness average per State
IBU_ABV$State= factor(IBU_ABV$State, level =IBU_ABV$State[order(IBU_ABV$IBU_mean)])
plot1=IBU_ABV %>% ggplot(aes(y=State, x=IBU_mean, color=State)) + geom_col(width=.5) +
  ggtitle("Bitternes average per State") + xlab("Bitternes average") + 
  ylab("State")+
  theme(axis.text = element_text(size = 9)) 

#ABV average per State
IBU_ABV$State= factor(IBU_ABV$State, level =IBU_ABV$State[order(IBU_ABV$ABV_mean)])
plot2=IBU_ABV %>% ggplot(aes(y=State, x=ABV_mean, color=State)) + geom_col(width=.5) +
  ggtitle("Alcohol average per State") + xlab("Alcohol by volume average") + 
  ylab("State") +
  theme(axis.text = element_text(size = 9)) 

```

```{r, fig.width=16,fig.height=7}
grid.arrange(plot1, plot2, ncol=2)
```

- The highest mean bitterness unit average state is Nevada while the lowest is Wisconsin
- State with the highest alcohol average per state is Nevada and lowest is UTAH


**Top 3 states with highest Bitterness**

```{r kable3, results = 'asis' }
kable(head(ibu_state %>% arrange(desc(IBU_mean)),3),"simple")
```

**States with Lowest Bitternes unit** \

```{r kable4, results = 'asis' }
kable(head(ibu_state %>% arrange(IBU_mean),3) ,"simple")
```

**Top states with highest Alcohol volume** \

```{r kable5, results = 'asis' }
#ABV mean per State
kable(head(abv_state %>% arrange(desc(ABV_mean)),3) ,"simple")
```

**States with Lowest Alcohol unit** \

```{r kable6, results = 'asis' }
kable(head(abv_state %>% arrange((ABV_mean)),3) ,"simple")
```

# Alcohol by volume distribution 
```{r include=FALSE}
#Histogram
plot4=all_data %>% ggplot(aes(x=ABV)) + geom_histogram(color="#FF9933", fill="white") +
  ggtitle("Distribution of alcohol by volume average") + 
  xlab("Alcohol by volume average") +
  ylab("Count") + 
  geom_vline(aes(xintercept=mean(ABV,na.rm=T)),
             color="#D55E00", linetype="dashed", size=1)


#Alcohol volume distribution boxplot
plot3=all_data %>% ggplot(aes(y=ABV)) + geom_boxplot(color="#FF9933") +
  ggtitle("Distribution of alcohol by volume average") + ylab("Alcohol by volume average") 

#ggplotly(plot3)

```
```{r, fig.width=16,fig.height=7}
grid.arrange(plot3, plot4, ncol=2)
```
```{r include=FALSE}
#Get more information on the outliers 
all_data %>% filter(ABV<0.02) #Low end outliers
top_out=(all_data %>% filter(ABV>0.067)) #High end outliers
top_out=top_out %>% group_by(State) %>% summarize(per_state=n()) %>% arrange(desc(per_state))
```

- The distribution of Alcohol by volume average is right skewed 
- There are outlier on both edges.
- There is only 1 outlier beneath the lower fence: 0.001 in California. That is a beer with very low Alcohol volume, should verify if the value is correct. The brewery was confirmed but no information on the beer. 
- The outer fence is 0.067 but the max is 0.12. Above the upper fence there are 585 outliers.
- The 4 top states with most outliers are within the the top states with highest mean Alcohol volume : 
    - 82 from Colorado which is the 6th State with highest alcohol mean, \
    - 64 from California which is the 11th state with highest mean Alcohol Volume \
    - 46 from Michigan which is the 5th state with highest mean Alcohol Volume \
    - 43 from Indiana which is the 4th highest mean Alcohol Volume \

- Alcohol vs Bitterness scatterplot \

```{r}
all_data %>% 
ggplot(aes(x= ABV, y=IBU)) + 
geom_point(color="#FF33CC") +
  ggtitle("Alcohol vs Bitterness") + xlab("Alcohol by volume") +
  ylab("International bitterness unit")  + theme(plot.title = element_text(hjust = 0.5)) 

```

- Alcohol vs Bitterness box plot divided in segments for better visualization 


```{r}
all_data %>% 
ggplot(aes(x= ABV, y=IBU, color=as.factor(cut_width(all_data$ABV,.009)))) + 
geom_boxplot(mapping = aes(group = cut_width(ABV,.009))) +
  ggtitle("Alcohol vs Bitterness") + xlab("Alcohol by volume") +
  ylab("International bitterness unit") +
  scale_colour_discrete(name="Alcohol Segment")+ 
theme_fivethirtyeight() +
  theme(plot.title = element_text(hjust = 0.5)) 
```



- Performing a cor.test to check correlation  \
```{r include=TRUE}
# Shapiro-Wilk normality test for both IBU and ABV
#shapiro.test(all_data$IBU) # => p = 2.2e-16
#shapiro.test(all_data$ABV) # => p = 2.2e-16

#Since neither have a normal distribution we are going to use spearman method
cor.test(all_data$IBU,all_data$ABV,method = "spearman") # Finding correlation
```
# Is there a relationship between Alcohol and bitterness? \


- The EDA Analysis suggest a relationship between alcohol and bitterness \
- The mean per segment continues to show an increasing pattern after missing values were imputed \
- The correlation between alcohol volume per beer and bitterness per unit beer per pairplot is 78%. It increased by 10 percent after the missing values were addressed \
- It coincides with the idea that the less alcohol, the less bitter and that the more alcohol, the more bitter \
- The correlation test rejected the null hypothesis that r=0 with a p-value of 2.2e-16^. Suggesting that there is a significant correlation between bitterness unit and alcohol volume. It provided an estimate correlation coefficient value of .78 \



```{r include=TRUE}

#Normalize IBU and ABV for model
#Min-Max normalization function
min_max_norm <- function(x) {
    return((x - min(x)) / (max(x) - min(x)))
}

all_norm =  as.data.frame(lapply(all_data[7:8], min_max_norm))
all_data["ABV_norm"]=all_norm[1] # adding the normalized_ABV column to all the data
all_data["IBU_norm"]=all_norm[2] # adding the normalized_IBU column to all the data


#Data frame with both IPA and the rest of the ALE
ale_ipa=all_data[str_detect(all_data$Style,regex("[:alpha:]*( Ale)[:alpha:]*|(IPA)",ignore_case = TRUE)),]
a=dim(ale_ipa) #1492 rows with Ale and IPA
m=dim(all_data) #2348

b=length(unique(ale_ipa$Style)) #33 beers with Ale and IPA
s=length(unique(all_data$Style)) #100 Style beers
beers_excluded=setdiff(unique(all_data$Style),unique(ale_ipa$Style)) #Style of beer excluded

#dim(ale_ipa) #length of IPA and Ale data 1492
#dim(all_data) #all data was 2347 , 832 rows were dropped without labels

#Label
ale_ipa$IPAorAle=ifelse(str_detect(ale_ipa$Style,regex("[:alpha:]*(IPA)[:alpha:]*",ignore_case = TRUE)),"IPA","Ale")

```

# KNN model
- After finding the best K
```{r include=FALSE}

# KNN Loop to find best K values for the model
# Loop for many k and the average of many training / test partition
set.seed(5)
numks = 60
masterAcc = matrix(nrow = 100, ncol = numks)

#Finding best K  
for(j in 1:100)
{
accs = data.frame(accuracy = numeric(60), k = numeric(60))
trainIndices = sample(1:dim(ale_ipa)[1],round(.7 * dim(ale_ipa)[1]))
train = ale_ipa[trainIndices,]
test = ale_ipa[-trainIndices,]
for(i in 1:numks)
{
  classifications = knn(train[,c(12,13)],test[,c(12,13)],as.factor(train$IPAorAle), prob = TRUE, k = i)
  table(as.factor(test$IPAorAle),classifications)
  CM = confusionMatrix(table(as.factor(test$IPAorAle),classifications))
  masterAcc[j,i] = CM$overall[1]
}
}

MeanAcc = colMeans(masterAcc)

#The best K
which.max(MeanAcc)

```

# KNN was performed with a K of 5
```{r}

trainIndices = sample(1:dim(ale_ipa)[1],round(.7 * dim(ale_ipa)[1]))
train = ale_ipa[trainIndices,]
test = ale_ipa[-trainIndices,]
classifications = knn(train[,c(12,13)],test[,c(12,13)],as.factor(train$IPAorAle), prob = TRUE, k = 5)
CM = confusionMatrix(table(as.factor(test$IPAorAle),classifications))
```
```{r kable7, results = 'asis' }
#ABV mean per State
kable(table(as.factor(test$IPAorAle),classifications) ,"simple")
```


Results:
- 84% Accuracy  \
- 90% Sensitivity  \
- 76% Specificiy \


## Cross validation with KNN
```{r} 
classifications = knn.cv(ale_ipa[,c(12,13)],as.factor(ale_ipa$IPAorAle), prob = TRUE,k = 5)
CM=confusionMatrix(classifications,as.factor(ale_ipa$IPAorAle))
```
```{r kable8, results = 'asis' }
#ABV mean per State
kable(table(as.factor(ale_ipa$IPAorAle),classifications) ,"simple")
```

# KNN scores using cross validation: \
- 82% Accuracy 
- 87% Sensitivity 
- 75% Specificiy


## NaybeBayes Model
```{r}
model = naiveBayes(train[,c(12,13)],train$IPAorAle)
CM = confusionMatrix(table(predict(model,test[,c(12,13)]),test$IPAorAle))
```
```{r kable9, results = 'asis' }
#ABV mean per State
kable(table(predict(model,test[,c(12,13)]),test$IPAorAle))
```


Naive Bayes score:
- 79% accuracy 
- 83% Sensitivity 
- 73% Specificity 

***Per the performance of the three models, (KNN, KNN with Cross validation and Naybe bayes), utilizing both IBU and ABV, suggest that there is a relationship between Alcohol volume and Bitterness unit between IPA beers and the rest of the Ale beers. The models had an accuracy range of 74% to 84%, with a sensitivity range of 85%-87% and a specificity of 74%-76%. To only be using IBU and ABV as the data to predict the type of beer it did a really good job. It showed that the two different style of beers have a specific range of Bitterness + Alcohol. At the same time it reinforces that there is a relationship between IBU and ABV. If there was no correlation between both IBU and ABV KNN would have not been able to create the groups with good scores.***


# Finding the mean ABV and IBU for styles of beer selected
```{r}
ale_ipa %>% group_by(IPAorAle) %>% summarize(mean_ibu=mean(IBU),mean_ABV=mean(ABV))
```


***Indian Pale Ale style of beer has a higher Bitter and Alcohol unit than that of the rest of the Ale beers***

# Map to visualize the top style beer per state and its quantity
```{r, fig.width=16,fig.height=7}
#get maps data
states=map_data('state')

#Make data frame with abbreviation and state name
st_abb=data.frame(abb=state.abb, region=tolower(state.name))

#creating a data frame with the top beer and quanity
#grouping per state and beer label
ipa_ale1=ale_ipa %>% group_by(State, IPAorAle) %>% tally()
#creating a column per label
ipa_ale1=spread(ipa_ale1, IPAorAle, n)
#Replacing with 0 states that don't have ABV or IBU and resulted in NA
ipa_ale1[is.na(ipa_ale1)]=0
#creating column the with selecting the style most popular in that state
ipa_ale1$mayority=ifelse(ipa_ale1$Ale>ipa_ale1$IPA,"-ale", "IPA")
#creating a column with the number of beers of the most popular style
ipa_ale1$mayority_val=ifelse(ipa_ale1$Ale>ipa_ale1$IPA, ipa_ale1$Ale, ipa_ale1$IPA)
#replacing those that have the same 
ipa_ale1$mayority[ipa_ale1$Ale==ipa_ale1$IPA] <- "IPA and Ale"
ipa_ale1$mayority_val[ipa_ale1$Ale==ipa_ale1$IPA] <- 0
#dim(ale_ipa[ale_ipa$IPAorAle=="IPA",]) #560
#dim(ale_ipa[ale_ipa$IPAorAle=="Ale",]) #932
brew=ipa_ale1 %>% select(-Ale, -IPA)
brew$State <- trimws(brew$State, which = c("both"))

#merging top beers and quantity with the map data
#change name to abb 
colnames(brew)[1]="abb"
#setdiff(unique(st_abb$abb),unique(brew$abb)) 
#Merge count in data with the region information
#Using new region column to combine with states data
brew=merge(brew,st_abb,by="abb")
#Using new region column to combine with states data
geo= merge(states,brew, by="region", all.x=TRUE)
#geo=geo %>% arrange(order)
geo=geo[order(geo$order),]
center = data.frame(region=tolower(state.name), long=state.center$x, lat=state.center$y)
center= merge(brew, center, by="region", all.x=TRUE)



ggplot(geo, aes(x=long, y=lat, color=as.factor(mayority))) +
  geom_polygon(aes(group=group, fill=mayority_val)) +
  geom_text(data=center, aes(long, lat, label=mayority_val), color="black") +
  ggtitle("Favorite beer per state and its quantity") +
  coord_map()  +
  theme(title=element_text(size=20)) +
  ylab("Latitude")+ xlab("Longitude") + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_gradient(low="white",
                      high="violetred3",
                      guide="colorbar", name="Number of breweries of \n the popular beer type")+
  scale_color_manual(values=c("white", "lightseagreen", "hotpink4"), name="Beer type") +
  guides(color = guide_legend(override.aes = list(size = 5) )) + 
  theme(legend.background = element_rect(fill = "snow2"),legend.text=element_text(size=10),legend.title=element_text(size=13),plot.title = element_text(size=22)) 


```
- Most states have more –ale beers than India Pale ale
- 4 states have the same amount of IPA beers and the rest of the –ale beers
- 4 states in the north-east have more IPA beers than –ale beers
